Determinism, Complexity, and Predictability in Computer Performance

### 2014-07-30 model.tex
pushing changes midstream, in case of lossage

These four forecast methods were chosen to be a representative
sampling of standard prediction strategies, but they do not, of
course, cover that space exhaustively.  Our goal here is an empirical
assessment of the relationship between predictability and complexity,
not formal results about a ``best'' predictor for a given time series.
From a practitioner's standpoint, it would be useful to know \emph{ a
  priori} if a time series contains enough predictive structure to
make it worth spending the time and effort searching for a good
forecast method.  It would also be useful to know if a given method is
inadequate---that is, if an alternative method could do better and so
one should continue searching.  \alert{[[need to check that this is not
    redundant, given the other text in this section (and elsewhere in
    the paper).  it may be better to put this in the conclusion.]]}

@@

To study the relationship between predictability and complexity, we
use the four methods outlined above to generate predictions of all 360
traces described in Section~\ref{sec:methods}, then calculate the
error of the predictions with respect to the true continuations.
Specifically, we split each time series into two pieces: the first
90\%, referred to as the ``initial training" signal and denoted
$\{x_i\}_{i=1}^{n}$, and the last 10\%, known as the ``test" signal
$\{c_j\}_{j=n+1}^{k+n+1}$.  The initial training signal is used to
build the model, following the procedures described in the previous
section; that model is used to generate a prediction of the value of
$x_{n+1}$, which is then compared to the true continuation, $c_{n+1}$.
The model is then rebuilt using $\{x_i\}_{i=1}^{n+1}$ and the process
repeats $k$ times, out to the end of the observed time series.  This
``one step prediction'' process is not technically necessary in the
LMA method, whose ability to generate accurate predictions is limited
only by the positive Lyapunov exponents of the system.  However, the
performance of the other three methods used here will degrade severely
if the associated models are not periodically rebuilt.  In order to
make the comparison fair, we used an iterative one-step prediction
schema \emph{for all four methods}.  This has the slightly confusing
effect of causing the ``test'' signal to be used both to assess the
accuracy of each model and for periodic refitting.

@@

As a numerical measure of prediction accurary, we calculate the mean
absolute scaled error (MASE) between the true and predicted signals:
%
$$MASE = \sum_{j=n+1}^{k+n+1}\frac{|p_j-c_j|
}{\frac{k}{n-1}\sum^n_{i=2}|x_{i}-x_{i-1}|}$$
%
This error metric was introduced in \cite{MASE} as a ``generally
applicable measurement of forecast accuracy without the problems seen
in the other measurements."  Like many error metrics, MASE is a
normalized measure: the scaling term in the denominator
%
% of the MASE value
% $$\frac{1}{n-1}\sum^n_{i=2}|X_{i,obs}-X_{i-1,obs}|$$
%
is the average in-sample forecast error for a random-walk prediction
over the initial training signal $\{x_i\}^n_{i=1}$.  That is, $MASE<1$
means that the prediction error in question was, on the average,
smaller than the error of a random-walk forecast on the same data.
Analogously, $MASE>1$ means that the corresponding prediction method
did \emph{worse}, on average, than the random-walk method.  While its
comparative nature is somewhat different than traditional metrics like
normalized root mean squared error, MASE has the significant advantage
of allowing one to make a fair comparison across varying methods,
prediction horizons, and signal scales---attributes that are key to
any broad study of predictability.

@@

  \hline\hline
  \end{tabular*}
  \end{center}
 \label{tab:error}
  \end{table*}%
In view of the discussion at the end of the previous paragraph, the
fact that the values in the second-from-right column are not
identically equal to 1.00 may be somewhat surprising.  \alert{[[Need
      to add a sentence here...presumably it's because not all RW
      forecasts are the same?  Why is this more of an issue in
      \svdthree?]]}  Comparing the values in Table~\ref{tab:error}
with the geometry of the plots in Figure~\ref{fig:forecast-example},
one can see some obvious correspondences.  The average LMA MASE score
for the \col signals was $0.050$, for instance, while \arima scored
much worse ($0.599$).  That is, LMA performed roughly 20 times better
on \col signals than a random-walk predictor, while \arima only
outperformed random walk by a factor of 1.7.  This is in accordance
with the visual appearance of the corresponding images in
Figure~\ref{fig:forecast-example}.  While the correspondence between
MASE score and the visual appearance of these kinds of plots is clear
cut in this case, that is not always so.  The plots of LMA predictions
of \col and \svdfive both lie near the diagonal, for instance, but the
corresponding MASE scores were very different: $0.050$ for \col and
$0.718$ for \svdfive (resp., 20 and 1.4 times better than random-walk
forecasts of the same signals).  This is a result of the normalization
in the MASE score calculation.  Recall from Section~\ref{sec:simple}
that the random-walk method performs especially poorly on signals that
oscillate rapidly.  \col fits this description, so the random-walk
error on this signal---the denominator of the four MASE scores in the
first row of the Table---is pathologically high, which skews those
scores down.  Random-walk prediction is very effective for the
\svdfive signal, on the other hand, so the denominator is small and
the MASE scores are skewed upwards.  Normalization is, as is often the
case, a double-edged sword.  Here, it facilitates comparisons across
signals of different types and lengths, but its particular pathologies
must be taken into account, as discussed at more length in
Section~\ref{sec:results}.

>>>>

These four forecast methods were chosen to be a representative
sampling of standard prediction strategies, but they do not, of
course, cover that space exhaustively.  Our goal here is an empirical
assessment of the relationship between predictability and complexity,
not formal results about a ``best'' predictor for a given time series.
From a practitioner's standpoint, it would be useful to know \emph{ a
  priori} if a time series contains enough predictive structure to
make it worth spending the time and effort searching for a good
forecast method.  It would also be useful to know if a given method is
inadequate---that is, if an alternative method could do better and so
one should continue searching.  \alert{need to check that this is not
  redundant, given the other text in this section (and elsewhere in
  the paper).  it may be better to put this in the conclusion.}

@@

To study the relationship between predictability and complexity, we
use the four methods outlined above to generate predictions of all 120
\alert{we had 360 there.  seems like it should be 8 times 15 which is
  120.}  traces described in Section~\ref{sec:methods}, then calculate
the error of the predictions with respect to the true continuations.
Specifically, we split each time series into two pieces: the first
90\%, referred to as the ``initial training" signal and denoted
$\{x_i\}_{i=1}^{n}$, and the last 10\%, known as the ``test" signal
$\{c_j\}_{j=n+1}^{k+n+1}$.  The initial training signal is used to
build the model, following the procedures described in the previous
section; that model is used to generate a prediction of the value of
$x_{n+1}$, which is then compared to the true continuation, $c_{n+1}$.
The model is then rebuilt using $\{x_i\}_{i=1}^{n+1}$ and the process
repeats $k$ times, out to the end of the observed time series.  This
``one step prediction'' process is not technically necessary in the
LMA method, whose ability to generate accurate predictions is limited
only by the positive Lyapunov exponents of the system.  However, the
performance of the other three methods used here will degrade severely
if the associated models are not periodically rebuilt.  In order to
make the comparison fair, we used an iterative one-step prediction
schema \emph{for all four methods}.  This has the slightly confusing
effect of causing the ``test'' signal to be used both to assess the
accuracy of each model and for periodic refitting.

@@

As a numerical measure of prediction accurary, we calculate the mean
absolute scaled error (MASE) between the true and predicted signals:
%
$$MASE = \sum_{j=n+1}^{k+n+1}\frac{|p_j-c_j|
}{\frac{k}{n-1}\sum^n_{i=2}|x_{i}-x_{i-1}|}$$
%
This error metric was introduced in \cite{MASE} as a ``generally
applicable measurement of forecast accuracy without the problems seen
in the other measurements."  Like many error metrics, MASE is a
normalized measure: the scaling term in the denominator
%
% of the MASE value
% $$\frac{1}{n-1}\sum^n_{i=2}|X_{i,obs}-X_{i-1,obs}|$$
%
is the average in-sample forecast error for a random-walk prediction
over the initial training signal $\{x_i\}^n_{i=1}$.  That is, MASE$<1$
means that the prediction error in question was, on the average,
smaller than the average error of a random-walk forecast on the
training data.  Analogously, MASE$>1$ means that the corresponding
prediction method did \emph{worse}, on average, than the random-walk
method.  While its comparative nature is somewhat different than
traditional metrics like normalized root mean squared error, MASE has
the significant advantage of allowing one to make a fair comparison
across varying methods, prediction horizons, and signal
scales---attributes that are key to any broad study of predictability.

@@

  \hline\hline
  \end{tabular*}
  \end{center}
 \label{tab:error}
  \end{table*}%
In view of the discussion at the end of the previous paragraph, the
fact that the values in the second-from-\alert{left} column are not
identically equal to 1.00 may be somewhat surprising.  This can happen
due to differences between in-sample and out-of-sample forecasting;
the last 10\% of \svdthree, for instance---the out-of-sample
signal---was easier, on the average, for RW to predict than the first
90\%. \alert{[[Liz to do: add pointer to Josh's paragraph for deeper
      analysis]]}.

Comparing the values in Table~\ref{tab:error} with the geometry of the
plots in Figure~\ref{fig:forecast-example}, one can see some obvious
correspondences.  The average LMA MASE score for the \col signals was
$0.050$, for instance, while \arima scored much worse ($0.599$).  That
is, LMA performed roughly 20 times better on \col signals than a
random-walk predictor, while \arima only outperformed random walk by a
factor of 1.7.  This is in accordance with the visual appearance of
the corresponding images in Figure~\ref{fig:forecast-example}.

In other cases, the correspondence between MASE score and the visual
appearance of these kinds of plots is not so clear cut.  The plots of
LMA predictions of \col and \svdfive both lie near the diagonal, for
instance, but the corresponding MASE scores were very different:
$0.050$ for \col and $0.718$ for \svdfive (resp., 20 and 1.4 times
better than random-walk forecasts of the same signals).  This is a
result of the normalization in the MASE score calculation.  Recall
from Section~\ref{sec:simple} that the random-walk method performs
especially poorly on signals that oscillate rapidly.  \col fits this
description, so the random-walk error on this signal---the denominator
of the four MASE scores in the first row of the Table---is
pathologically high, which skews those scores down.  Random-walk
prediction is very effective for the \svdfive signal, on the other
hand, so the denominator is small and the MASE scores are skewed
upwards.

Normalization is, as is often the case, a double-edged sword.  Here,
it facilitates comparisons across signals of different types and
lengths, but its particular pathologies must be taken into account, as
discussed at more length in Section~\ref{sec:results}.

### 2014-07-30 results.tex
pushing changes midstream, in case of lossage

\alert{Are there any updates that we need to do besides the curve
  shape and the addition of random walk results?}

@@

The experiments below involve three different prediction methods
applied to time-series data from eight different systems: {\tt
  col\_major}, {\tt 403.gcc}, and the six different regimes of the
{\tt dgesdd} signal in Figure~\ref{fig:svd-ts-colored}.  The objective
of these experiments was to explore how prediction accuracy is related
to WPE.
%
% , and how that relationship depends on the generating process
% and the prediction method.
%
Working from the first 90\% of each signal, we generated a prediction
of the last 10\% using the \naive, \arima, and LMA prediction methods,
as described in Section~\ref{sec:accuracy}, then calculated the MASE
value of those predictions.  We also calculated the WPE of each time
series using a wordlength chosen via the procedure described in
Section~\ref{sec:meaComplex}.  In order to assess the run-to-run
variability of these results, we repeated all of these calculations on
15 separate trials: i.e., 15 different runs of each program.

@@

Figure~\ref{fig:wpe_vs_mase_best} plots the WPE values versus the
corresponding MASE values of the best prediction of each of the
\alert{[[108?]]} time series in this study
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\columnwidth]{figs/new_prediction_vs_entropy}
  \caption{Weighted permutation entropy vs. mean absolute scaled error
    (MASE) of the best prediction of each time series.
%
% (across all trials and all prediction methods)
%
The solid curve is a least-squares log fit of these points.
% 
% of all the points except those from {\tt dgesdd$_1$}, which we have
% excluded for reasons explained in the text.
% 
The dashed curves reflect the standard deviation of the model in its
parameter space.  The meaning of the shaded region is explained in the
text.}
  \label{fig:wpe_vs_mase_best}
\end{figure}
\alert{As the plot shows, the best-case prediction error for each trace from
these eight systems is roughly proportional to the log of the weighted
permutation entropy, which is consistent with our first conjecture.}
The dashed line in the figure, a least-squares fit of $y = a \log(b x
+ 1)$ to these errors,
% 
% \footnote{with the exception of {\tt
%     dgesdd$_1$}, for reasons described below} 
% 
captures this relationship between WPE and prediction error quite
nicely.  The logarithmic shape of this curve makes sense here because
any signal with zero entropy should be perfectly predictable, and so
should ideally have a MASE value of zero.  The shaded region shows the
standard deviation of the model in its parameter space: i.e., $y = a
\log(b x + 1) + c x^{16}$ with $\pm$ one standard deviation on each of
the three parameters.

This is not a formal result.  The forecast methods and data sets used
here were chosen to span the space of standard prediction strategies
and the range of dynamical behaviors, but they do not cover those
spaces exhaustively.  Our goal here is an \emph{empirical} assessment
of the relationship between predictability and complexity, not formal
results about a ``best'' predictor for a given time series.  There may
be other methods that produce lower MASE values than those in
Figure~\ref{fig:wpe_vs_mase_best}, but the sparseness of the points
above and below the curve in this plot strongly suggests a pattern of
(inverse) correlation between the underlying predictability of a time
series and its WPE.  The rest of this section describes our results in
more detail---including the measures taken to assure meaningful
comparisons across methods, trials, and programs---and elaborates on
the meaning of the dashed curve in the figure.

Figure~\ref{fig:wpe_vs_mase_all} shows WPE vs. MASE plots for the full
set of experiments.
\begin{figure}
  \centering
%    \includegraphics[width=\columnwidth]{figs/ARIMA_prediction_vs_entropy}
%    \includegraphics[width=\columnwidth]{figs/ARIMA_prediction_vs_entropy}
%    \includegraphics[width=\columnwidth]{figs/LMA_prediction_vs_entropy}
  \includegraphics[width=\columnwidth]{figs/new_predictions_vs_entropy4a}
\caption{WPE vs. MASE for all trials, methods, and systems---with the
  exception of \svdone, \svdthree, and \svdfive are omitted from the
  top plot for scale reasons. \alert{there's also a version with all plots
  stacked (4x1) instead of (2x2)}
%
% ; their MASE scores are $2.676 \pm 4.328$, $31.39 \pm 0.282$, and
%   $20.87 \pm 0.192$ respectively.
%
Numerical values, including means and standard deviations of the
errors, can be found in Table~\ref{tab:error}.  The dashed curve and
shaded region is the same as in the previous figure.  }
    \label{fig:wpe_vs_mase_all}
\end{figure}
There are 15 points in each cluster, one for each trial.  (The points
in Figure~\ref{fig:wpe_vs_mase_best} are the leftmost of the points
for the corresponding trace in any of the three plots in
Figure~\ref{fig:wpe_vs_mase_all}.)  The WPE values do not vary very
much across trials.  For most traces, the variance in MASE scores is
low as well, resulting in small, tight clusters.  In some
cases---\arima predictions of {\tt col\_major}, for instance---the
MASE variance is larger, which spreads out the clusters horizontally.
The distribution of the cluster \emph{positions} is quite different in
the three plots.  The mean MASE scores of predictions generated with
the nonlinear LMA method are generally closer to the dashed curve; the
\arima method clusters are more widely spread, the \naive ~clusters
even more so.  This geometric validation of the second conjecture in
this paper is discussed in more detail in the following paragraphs.

@@

The WPE of \gcc is higher still: $0.943 \pm 0.001$.  This system
transmits very little information forward in time and provides almost
no structure for prediction methods to work with.  \alert{This needs to be reworded to reflect the typo, \naive is no longer the winner. MASE is $1.797 \pm 0.010$, rw actually wins $MASE = 1.1381 \pm 0.011$  Here, the \naive
~method performs best ($MASE=0.951 \pm 0.001$), simply because its
calculation of the mean filters out the noise in the signal.  Since
fitting a hyperplane using least squares should have similar effects,
the fact that LMA outperforms \arima ($1.530 \pm 0.021$ vs. $1.837
\pm 0.016$) is somewhat counterintuitive.  However, the small amount
of predictive structure that is present in this signal is nonlinear
(cf.,~\cite{mytkowicz09}), and LMA is designed to capture and exploit
that kind of structure.}

\svdone---the dark blue segment of
Figure~\ref{fig:svd-ts-colored}---behaves very differently than the
other seven systems in this study.  Though its weighted permutation
entropy is very high ($0.957 \pm 0.016$), two of the three prediction
methods do quite well on this signal, yielding MASE scores of $0.827
\pm 0.076$ (LMA) and $0.714 \pm 0.075$ (\arima).  This pushes the
corresponding clusters of points in the bottom two images in
Figure~\ref{fig:wpe_vs_mase_all} well above the trend followed by the
other seven signals.  The MASE scores of the predictions that were
produced by the \naive ~method for this system, however, are highly
inconsistent.  The majority of the blue diamond-shaped points on the
top plot in Figure~\ref{fig:wpe_vs_mase_all} are clustered near a MASE
score of 0.6, which is better than LMA and \arima.  In five of the
15 \svdone trials, however, there were step changes in the signal.
The \naive ~method has a very difficult time with signals like this,
particularly if there are multiple step changes.  This raised the MASE
scores of these trials, pushing the corresponding points to the
right\footnote{This includes the cluster of three points near MASE
  $\approx 2.5$, as well as two points that are beyond the domain of
  the graph, at MASE 11.2 and 14.8.}, and in turn raising both the
mean and variance of this set of trials.

>>>>

@@

The experiments below involve four different prediction methods
applied to time-series data from eight different systems: {\tt
  col\_major}, {\tt 403.gcc}, and the six different segments of the
{\tt dgesdd} signal in Figure~\ref{fig:svd-ts-colored}.  The objective
of these experiments was to explore how prediction accuracy is related
to WPE.
%
% , and how that relationship depends on the generating process
% and the prediction method.
%
Working from the first 90\% of each signal, we generated a prediction
of the last 10\% using the \naive, \arima, and LMA prediction methods,
as described in Section~\ref{sec:accuracy}, then calculated the MASE
value of those predictions.  We also calculated the WPE of each time
series using a wordlength chosen via the procedure described in
Section~\ref{sec:meaComplex}.  In order to assess the run-to-run
variability of these results, we repeated all of these calculations on
15 separate trials: i.e., 15 different runs of each program.

@@

Figure~\ref{fig:wpe_vs_mase_best} plots the WPE values versus the
corresponding MASE values of the \emph{best} prediction for each of
the 120 time series in this study.
\begin{figure}
  \centering
  \includegraphics[width=1.1\columnwidth]{figs/new_prediction_vs_entropy}
  \caption{Weighted permutation entropy vs. mean absolute scaled error
    (MASE) of the best prediction of each time series.  The solid
    curve is a least-squares log fit of these points.
% 
% of all the points except those from {\tt dgesdd$_1$}, which we have
% excluded for reasons explained in the text.
% 
The dashed curves reflect the standard deviation of the model in its
parameter space.  The meaning of the shaded region is explained in the
text.}
  \label{fig:wpe_vs_mase_best}
\end{figure}
% 
% As the plot shows, the best-case prediction error for each trace from
% these eight systems is roughly proportional to the log of the weighted
% permutation entropy, which is consistent with our first conjecture.
% 
There is an obvious upward trend, which is consistent with our first
conjecture.  However, a simple linear fit is a bad idea here.  First,
any signal with zero entropy should be perfectly predictable (i.e.,
MASE $\approx 0$), so any curve fitted to these data should pass
through the origin.  Moreover, WPE does not grow without bound, so one
would expect the patterns in the WPE-MASE pairs to reach some sort of
asymptote.  For these reasons, we chose to fit a function of the form
$y = a \log(b x + 1)$ to these points, with $y =$ WPE and $x=$ MASE.
\alert{can we cite a figure of merit for this fit?  we also need to
  say what $a$ and $b$ were.}  The solid curve in the figure shows
this fit; the dashed curves show the standard deviation of this model
in its parameter space: i.e., $y = a \log(b x + 1)$ with $\pm$ one
standard deviation on each of the two parameters.  Points that fall
within this deviation volume \alert{(light grey)} correspond to
predictions that are comparable to the best ones found in this study;
points that fall \emph{above} that volume \alert{(dark grey)} are
better still.  We chose to truncate the shaded region because of a
subtle point about the MASE of an ideal predictor, which should not be
larger than 1 unless the training and test signals are different.
This is discussed at more length below.
%[[make sure that's true]]

The curves and regions in Figure~\ref{fig:wpe_vs_mase_best} are a
graphical representation of the correlation in the first finding.  It
is, we believe, a useful heuristic for determining whether a given
prediction method is well matched to a particular time series.  It is
not, of course, a formal result.  The forecast methods and data sets
used here were chosen to span the space of standard prediction
strategies and the range of dynamical behaviors, but they do not cover
those spaces exhaustively.  Our goal here is an \emph{empirical}
assessment of the relationship between predictability and complexity,
not formal results about a ``best'' predictor for a given time series.
There may be other methods that produce lower MASE values than those
in Figure~\ref{fig:wpe_vs_mase_best}, but the sparseness of the points
above and below the one-$\sigma$ region about the dashed curve in this
plot strongly suggests a pattern of correlation between the underlying
predictability of a time series and its WPE.  The rest of this section
describes these results and claims in more detail---including the
measures taken to assure meaningful comparisons across methods,
trials, and programs---and elaborates on the meaning of the different
curves and limits in the figure.

\alert{Needs to get added somewhere here: The positions of the WPE
  vs. error points, and the reasons to choose a particular function to
  fit to them, will be different if you use another error metric.
  This one is just good for error $=$ MASE.  But MASE is good for
  global comparisons and we got lots of other time-series data to fall
  on it (list), so we believe it.}

Figure~\ref{fig:wpe_vs_mase_all} shows WPE vs. MASE plots for the
\emph{full} set of experiments.
\begin{figure*}
  \centering
%    \includegraphics[width=\columnwidth]{figs/ARIMA_prediction_vs_entropy}
%    \includegraphics[width=\columnwidth]{figs/ARIMA_prediction_vs_entropy}
%    \includegraphics[width=\columnwidth]{figs/LMA_prediction_vs_entropy}
  \includegraphics[width=1.7\columnwidth]{figs/new_predictions_vs_entropy4a}
\caption{WPE vs. MASE for all trials, methods, and systems---with the
  exception of \svdone, \svdthree, and \svdfive are omitted from the
  top plot for scale reasons.
%
% ; their MASE scores are $2.676 \pm 4.328$, $31.39 \pm 0.282$, and
%   $20.87 \pm 0.192$ respectively.
%
Numerical values, including means and standard deviations of the
errors, can be found in Table~\ref{tab:error}.  The dashed curve and
shaded regions are the same as in the previous figure.  }
    \label{fig:wpe_vs_mase_all}
\end{figure*}
There are 15 points in each cluster, one for each trial.  (The points
in Figure~\ref{fig:wpe_vs_mase_best} are the leftmost of the points
for the corresponding trace in any of the three plots in
Figure~\ref{fig:wpe_vs_mase_all}.)  The WPE values do not vary very
much across trials.  For most traces, the variance in MASE scores is
low as well, resulting in small, tight clusters.  In some
cases---\arima predictions of {\tt col\_major}, for instance---the
MASE variance is larger, which spreads out the clusters horizontally.
The mean MASE scores of predictions generated with the nonlinear LMA
method, for instance, are generally closer to the dashed curve; the
\arima method clusters are more widely spread, the \naive ~clusters
even more so.

The main thing to note here, however, is not the geometric details of
the clusters, but the fact that many of them are to the right of
and/or below the dashed curve that identifies the boundary of the
shaded region.  That is, these predictions are not as good as our
heuristic suggests they could be.  Focusing in on any cluster of
points makes this clear: LMA works best for \svdsix, for instance,
followed by the random-walk prediction method, then \arima and \naive.
The idea behind the shaded region is to identify when there is
predictive structure in the data that a given method is not exploiting
(viz., \arima on \svdsix)---i.e., that one could do better with
another forecast strategy.

The distribution of the cluster \emph{positions} is quite different in
the three plots.

@@

The WPE of \gcc is higher still: $0.943 \pm 0.001$.  This system
transmits very little information forward in time and provides almost
no structure for prediction methods to work with.  \alert{This needs
  to be reworded to reflect the typo, \naive is no longer the
  winner. MASE is $1.797 \pm 0.010$, rw actually wins MASE$=1.1381 \pm
  0.011$ Here, the \naive ~method performs best (MASE$=0.951 \pm
  0.001$), simply because its calculation of the mean filters out the
  noise in the signal.  Since fitting a hyperplane using least squares
  should have similar effects, the fact that LMA outperforms \arima
  ($1.530 \pm 0.021$ vs. $1.837 \pm 0.016$) is somewhat
  counterintuitive.  However, the small amount of predictive structure
  that is present in this signal is nonlinear
  (cf.,~\cite{mytkowicz09}), and LMA is designed to capture and
  exploit that kind of structure.}

\svdone---the dark blue segment of
Figure~\ref{fig:svd-ts-colored}---behaves very differently than the
other seven systems in this study.  Though its weighted permutation
entropy is very high ($0.957 \pm 0.016$), two of the three prediction
methods do quite well on this signal, yielding MASE scores of $0.827
\pm 0.076$ (LMA) and $0.714 \pm 0.075$ (\arima).  This pushes the
corresponding clusters of points in the bottom two images in
Figure~\ref{fig:wpe_vs_mase_all} well above the trend followed by the
other seven signals.  The MASE scores of the predictions that were
produced by the \naive ~method for this system, however, are highly
inconsistent.  The majority of the blue diamond-shaped points on the
top plot in Figure~\ref{fig:wpe_vs_mase_all} are clustered near a MASE
score of 0.6, which is better than LMA and \arima.  In five of the 15
\svdone trials, however, there were step changes in the signal.  The
\naive ~method has a very difficult time with signals like this,
particularly if there are multiple step changes.  This raised the MASE
scores of these trials, pushing the corresponding points to the
right\footnote{This includes the cluster of three points near MASE
  $\approx 2.5$, as well as two points that are beyond the domain of
  the graph, at MASE $\approx 11.2--14.8$.}, and in turn raising both
the mean and variance of this set of trials.
